DeepSphere: a graph-based spherical CNN

### 2019-11-11 rebuttal.md

# ICLR'20 rebuttal

## General comment about revised manuscript

We uploaded an improved manuscript thanks to the reviewers' comments.
The following has been updated:
* Slightly (very slightly) different statement of theorem 3.1
* new proposition (now called prop 1) formalizing the relation between theorem 3.1 and rotation equivariance

@@

As a non-expert of spherical CNN, I don't understand clearly the gap between the result Theorem 3.1 and showing the rotation equivalence. It would be nice to add some counterexample (i.e., in what situation the proposed approach does not have rotational equivalence while Theorem 3.1 holds).

### Answer

We thank the reviewer for their time assessing our work and their constructive feedback.

We prepared a revised manuscript, to be uploaded shortly, containing a deeper theoretical discussion about the gap between the result Theorem 3.1 and rotation equivariance. In a small proposition, to be added after theorem 3.1, we precise mathematically the relationship between these two concepts. In short, we proved that, if theorem 3.1 holds, our graph Laplacian L commutes with any rotation operator R in the limit of infinite sampling (pointwise), i.e. LRf(x)-RLf(x) \to 0, thus answering the reviewer's concerns about this subject.

### 2019-11-11 main.tex
small update
Updates from Overleaf

\todo{
rebuttal:
\begin{itemize}
    \item answer reviewers and commenters [mdeff \& nath]
    \item update the proof [nath \& martino]
    \item better link between the proof and equivarian
    \item publish the code and link it here
	\item swap AR and TC on figure 1
	\item add somewhere: exploit symmetries to reduce sample cost
\end{itemize}
}

@@

\paragraph{Theoretical convergence.}
We extend the work of Belkin and Nyiogi to a sufficiently regular, deterministic sampling scheme. \mart{Here we insert the definition of the extended graph laplacian and of the rotation operator Rg} Following their setting, we define the extended graph Laplacian operator as a linear operator $L_n^t: L^{2}(\mathcal{M}) \rightarrow L^{2}(\mathcal{M})$ such that
	\begin{equation}
	\label{eq:Heat Kernel Graph Laplacian operator}
	 L_n^tf(y) := \frac{1}{n}\sum_{i=0}^{n-1} e^{ -\frac{\|x_i-y\|^2}{4t}} \left(f(y)-f(x_i)\right).
	\end{equation}
Note that we have the following relation $\b{L}_n^t \b{f} = T_\V L_n^t f$.
Given a sampling scheme $\V = \{x_0, \dots, x_{n-1}\}$, we also define $\sigma_i$ to be the patch of the surface of the sphere corresponding to $x_i$, $A_i$ its corresponding area and $d_i$ the radius of the smallest ball in $\R^3$ containing $\sigma_i$ and centered at $x_i$. Define $d^{(n)} := \max_{i=0, \dots, n}d_i$ and $A^{(n)}=\max_{i=0, \dots, n}A_i$. Given any rotation $g\in SO(3)$, we define the corresponding rotation operator $\mathcal R_g$ such that for any function $f$ on the sphere 
\begin{equation}\label{eq:rotation operator}
\mathcal R_g f(x) := f(g^{-1}(x)).
\end{equation}
\begin{theorem}
	For a sampling $\V = \{x_i\in\S^2\}_{i=0}^{n-1}$ of the sphere that is equi-area and such that $d^{(n)} \leq \frac{C}{n^\alpha}, \ \alpha\in (0,\linefrac{1}{2}]$, for all $f: \S^2 \rightarrow \R$ Lipschitz with respect to the Euclidean distance in $\R^3$, for all $y\in\S^2$, there exists a sequence $t_n = n^\beta$ such that 
% 	the rescaled Heat Kernel Graph Laplacian $\frac{|\S^2|}{4\pi t_n^2}\b {L}^t_n$ converges pointwise to the Laplace Beltrami operator on the sphere $\Delta{\S^2}$  
	\begin{equation*}
		\lim_{n\to\infty}\frac{1}{4\pi t_n^2} L_n^{t_n}f(y) = \frac{1}{|\S^2|}\triangle_{\S^2}f(y)
	\end{equation*} 
% 	for $n\to\infty$.
	\label{theo:pointwise convergence for a regular sampling}
\end{theorem}
As a direct implication of this theorem, $\b{L}$ converges pointwisely toward the Laplace-Beltrami operator ($\Delta_{\S^2}$) in the sense explained in \cite{belkin2007convergence}. \mart{here we insert the link between consistency and equivariance} This operator commutes with rotation and hence this convergence is a strong hint for equivariance, as formalized in the following proposition.
%%%%%%%%%%%%%%%%%%%%%%%  link between consistency and equivariance %%%%%%%%%%%%%%%%%%%%%%%%%
\begin{prop}\label{prop:equivariance}
Under the hypothesis of theorem \ref{theo:pointwise convergence for a regular sampling}, we have the following pointwise convergence
\begin{equation*}
\forall x\in\mathbb S^2\quad\left| \mathcal R_g L_n^{t_n} f (x) - L_n^{t_n} \mathcal R_g f(x) \right| \xrightarrow{n\to\infty}0.
\end{equation*}
\end{prop}
%%%%%%%%%%%%%%%%%%%%%%% end of link between consistency and equivariance %%%%%%%%%%%%%%%%%%%%%%%%%

@@

		\caption{Normalized mean equivariance error \eqref{eq:mean equivariance error}. The number of pixels is $4s^2$ for the equiangular sampling and $12s^2$ for the Healpix sampling.}
		\label{fig:equivariance_error}
	\end{minipage}
	\hfill
	\begin{minipage}{0.35\linewidth}
		\centering
		\includegraphics[width=\linewidth]{kernel_widths}
		\caption{Optimal kernel widths.}
		\label{fig:kernel_widths}
		\vspace{1em}
		\includegraphics[height=4.5em]{lamp_000018}
		\hfill
		\includegraphics[height=4.5em]{lamp_000018_sphere_nobar}
		\caption{3D object represented as a spherical depth map.}
		\label{fig:depthmap}
		\vspace{1em}
		\includegraphics[width=\linewidth]{spectrum}
		\caption{Power spectral densities.}
		\label{fig:spectrum}
	\end{minipage}
	\vspace{-0.5cm}
\end{figure}

@@

%We thank Pierre Vandergheynst for advices, and Andreas Loukas for having processed the GHCN data.
%We thank Taco Cohen for his inputs on the intriguing results of our comparison with \citet{cohen2019gauge}.
%We thank the anonymous reviewers for their constructive feedback.
%We used the Python Graph Signal Processing package (PyGSP) \citep{pygsp} for computations and plots.
%The Python Graph Signal Processing package (PyGSP) \citep{pygsp} was used to build graphs, compute the Laplacian and Fourier basis, and perform graph convolutions.

>>>>

\todo{
rebuttal:
\begin{itemize}
    \item answer reviewers and commenters [mdeff \& nath]
    \item update the proof [nath \& martino]
    \item better link between the proof and equivarian
    \item publish the code and link it here
	\item swap AR and TC on figure 1
	\item add somewhere: exploit symmetries to reduce sample cost
	\item proposition 1 -> corollary 3.1?
\end{itemize}
}

@@

\paragraph{Theoretical convergence.}
We extend the work of Belkin and Nyiogi to a sufficiently regular, deterministic sampling scheme. \mart{Here we insert the definition of the extended graph laplacian and of the rotation operator Rg} Following their setting, we define the extended graph Laplacian operator as a linear operator $L_n^t: L^{2}(\mathcal{M}) \rightarrow L^{2}(\mathcal{M})$ such that
	\begin{equation}
	\label{eq:Heat Kernel Graph Laplacian operator}
	 L_n^tf(y) := \frac{1}{n}\sum_{i=0}^{n-1} e^{ -\frac{\|x_i-y\|^2}{4t}} \left(f(y)-f(x_i)\right).
	\end{equation}
Note that we have the following relation $\b{L}_n^t \b{f} = T_\V L_n^t f$.
Given a sampling scheme $\V = \{x_0, \dots, x_{n-1}\}$, we also define $\sigma_i$ to be the patch of the surface of the sphere corresponding to $x_i$, $A_i$ its corresponding area and $d_i$ the radius of the smallest ball in $\R^3$ containing $\sigma_i$ and centered at $x_i$. Define $d^{(n)} := \max_{i=0, \dots, n}d_i$ and $A^{(n)}=\max_{i=0, \dots, n}A_i$. Given any rotation $g\in SO(3)$, we define the corresponding rotation operator $\mathcal R_g$ such that for any function $f$ on the sphere 
\begin{equation}\label{eq:rotation operator}
\mathcal R_g f(x) := f(g^{-1}(x)).
\end{equation}
\begin{theorem}
	For a sampling $\V = \{x_i\in\S^2\}_{i=0}^{n-1}$ of the sphere that is equi-area and such that $d^{(n)} \leq \frac{C}{n^\alpha}, \ \alpha\in (0,\linefrac{1}{2}]$, for all $f: \S^2 \rightarrow \R$ Lipschitz with respect to the Euclidean distance in $\R^3$, for all $y\in\S^2$, there exists a sequence $t_n = n^\beta$, $\beta\in\mathbb R$ such that 
% 	the rescaled Heat Kernel Graph Laplacian $\frac{|\S^2|}{4\pi t_n^2}\b {L}^t_n$ converges pointwise to the Laplace Beltrami operator on the sphere $\Delta{\S^2}$  
	\begin{equation*}
		\lim_{n\to\infty}\frac{1}{4\pi t_n^2} L_n^{t_n}f(y) = \frac{1}{|\S^2|}\triangle_{\S^2}f(y)
	\end{equation*} 
% 	for $n\to\infty$.
	\label{theo:pointwise convergence for a regular sampling}
\end{theorem}
As a direct implication of this theorem, $\b{L}$ converges pointwise towards the Laplace-Beltrami operator ($\Delta_{\S^2}$) in the sense explained in \cite{belkin2007convergence}. \mart{here we insert the link between consistency and equivariance} This operator commutes with rotation and hence this convergence is a strong hint for equivariance, as formalized in the following proposition.
%%%%%%%%%%%%%%%%%%%%%%%  link between consistency and equivariance %%%%%%%%%%%%%%%%%%%%%%%%%
\begin{prop}\label{prop:equivariance}
Under the hypothesis of theorem \ref{theo:pointwise convergence for a regular sampling}, we have the following pointwise convergence
\begin{equation*}
\forall x\in\mathbb S^2\quad\left| \mathcal R_g L_n^{t_n} f (x) - L_n^{t_n} \mathcal R_g f(x) \right| \xrightarrow{n\to\infty}0.
\end{equation*}
\end{prop}
%%%%%%%%%%%%%%%%%%%%%%% end of link between consistency and equivariance %%%%%%%%%%%%%%%%%%%%%%%%%

@@

		\caption{Normalized mean equivariance error \eqref{eq:mean equivariance error}. The number of pixels is $n = 4 s^2$ for the equiangular sampling and $n = 12 s^2$ for the HEALPix sampling.}
		\label{fig:equivariance_error}
	\end{minipage}
	\hfill
	\begin{minipage}{0.35\linewidth}
		\centering
		\includegraphics[width=\linewidth]{kernel_widths}
		\caption{Optimal kernel widths.}
		\label{fig:kernel_widths}
		\vspace{1em}
		\includegraphics[height=4.5em]{lamp_000018}
		\hfill
		\includegraphics[height=4.5em]{lamp_000018_sphere_nobar}
		\caption{3D object represented as a spherical depth map.}
		\label{fig:depthmap}
		\vspace{1em}
		\includegraphics[width=\linewidth]{spectrum}
		\caption{Power spectral densities.}
		\label{fig:spectrum}
	\end{minipage}
	\vspace{-0.5cm}
\end{figure}

@@

%We thank Pierre Vandergheynst for advices, and Andreas Loukas for having processed the GHCN data.
%We thank Taco Cohen for his inputs on the intriguing results of our comparison with \citet{cohen2019gauge}.
%We thank the anonymous reviewers for their constructive feedback.
%The following packages were used for computation and plotting: pygsp, numpy, tensorflow, scipy, etc.
%We used the Python Graph Signal Processing package (PyGSP) \citep{pygsp} for computations and plots.
%The Python Graph Signal Processing package (PyGSP) \citep{pygsp} was used to build graphs, compute the Laplacian and Fourier basis, and perform graph convolutions.
